Dark matter halo shape at $z=0$ in the Auriga simulations: radial evolution and alignment with the stellar disk

### 2019-07-19 Document/paper.tex
edit

The first effect that we put in evidence in this paper is the effect
of baryons in producing rounder DM halos.
This effect has been already reported in other simulations \citep{Debattista08, Bryan13,
  Butsky16, Chua19, Artale19}.
The strength of the change depends on the numerical resolution, the
method to resolve the hydrodynamics and the models describing star
formation and stellar feedback.
The key concept behind these results is that the baryon distribution
influences and correlates with the dark matter halo shape. 
The broad tendency is that massive stellar disks correlate with
spherical dark matter distributions.

In order to explore this idea in the Auriga simulations we quantify
the  correlation between halo shape with baronyc disk properties. 
Looking into the measurements already reported by \cite{auriga} and
\cite{Pakmor17} we find three baryonic quantities that have the
strongest correlation with DM halo triaxiality: the central gas
density in a sphere of radius $1$kpc, the disk to total mass ratio and
the optical radius.

Figure \ref{fig:disk_correlations} summarizes the correlations of
these quantities with the triaxility at five diferent radii.
We use the Spearmen's rank correlation coefficient to quantify the
correlation strength.
This confirms our expectations. The strongest correlations are found
with the halo shape measured at radii smaller than $\sim 28$kpc, which
is close to the upper limit of the disk optical radius among our
simulation sample. 
The trend is such that halos with large triaxiality correlate with
high gas density and stellar disks with low mass and small size.
In turn, massive and large stellar disks with in a low density gas
environment correlate with low triaxiliaty.

Using the six halos in the high resolution sample we find that the
memory effect put foward by \cite{VeraCiro11}, that is the correlation
between the historical and radial evolution of the shape, only holds
for a fraction of the sample and is clearly broken for two halos in
the sample.
Given the small size of the sample we use to study this effect we
cannot find any significant correlation with other disk properties in
order to understand the shape amnesia. 
This effect has not been studied in detail in other simulations. 
We consider that understanding this effect is the first promising
extension of our studies to better understand dark matter halo shape
correlation with disk properties in Milky Way type galaxies.
avenues to continue a deeper study on dark matter halo shape.

The second promising extension of our results deals with the shape
alignment as a function of radius.
In concordance with previous results in DM only simulations we find
that the shells of halo shape are well aligned at different radii.
The surprising result we find is that in the presences of baryons
these shells twist as a function of radius.
\citep{Debattista08} found a similar effect, but only at radii smaller
than the disk radius. At radii larger than the disk radius, they find,
the dark matter shells stop twisting.
The twisting effect is robust across our sample of 30 halos. 
Also surprisingly at a radius of $0.25R_{200}\approx 56$ kpc the
alignment between the stellar disk and the dark matter halo is almost
perfect.

>>>>

The first effect that we put in evidence in this paper is the effect
of baryons in producing rounder DM halos.

The strength of the change depends on the numerical resolution, the
method to resolve the hydrodynamics and the models describing star
formation and stellar feedback \citep{Debattista08, Bryan13, Butsky16,
  Chua19, Artale19}.  
The key concept unifying these results is that the baryon distribution
influences and correlates with the dark matter halo shape. 
Here we find that the broad tendency is that massive stellar disks
correlate with spherical dark matter distributions.

In order to explore this idea in the Auriga simulations we quantify
the correlation between halo shape with baryonic disk properties. 
Looking into the measurements already reported by \cite{auriga} and
\cite{Pakmor17} we find three baryonic quantities that have the
strongest correlation with DM halo triaxiality: the central gas
density in a sphere of radius $1$kpc, the disk to total mass ratio and
the optical radius.

Figure \ref{fig:disk_correlations} shows the correlations of
these quantities with the triaxility at five diferent radii.
We use the Spearmen's rank correlation coefficient to quantify the
correlation strength.
We find that the strongest correlations are found with the halo shape
measured at radii smaller than $0.12R_{200}\sim 28$kpc, which 
is close to the upper limit of the disk optical radius among our
simulation sample. 
The trend is such that halos with large triaxiality correlate with
high gas density and stellar disks with low mass and small size.
In turn, massive and large stellar disks within a low density gas
environment correlate with low halo triaxiliaty.

Our second results deals with the shape alignment as a function of
radius. 
In concordance with previous results in DM only simulations we find
that the shells of halo shape are well aligned at different radii.
However, in the presences of baryons these shells twist as a function
of radius.
\citep{Debattista08} found a similar effect, but only at radii smaller
than the disk radius. At radii larger than the disk radius, they find,
the dark matter shells stop twisting.
In our case the he twisting effect is robust across our sample of 30
halos at all radii,
its strenght also correlates with the halo triaxiality for radii
smaller than $0.12R_{200}$. 
At larger radii 
Also surprisingly at a radius of $0.25R_{200}\approx 56$ kpc the
alignment between the stellar disk and the dark matter halo is almost
perfect.
